---
title: "Getting Started with OmopCausalGraph"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with OmopCausalGraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

```{r setup}
library(OmopCausalGraph)
```

## Overview

`OmopCausalGraph` lets you express a pharmacoepidemiology study as a causal DAG
where every node is anchored to an OMOP Standard Concept ID and can carry
one or more executable phenotype bindings.

## 1 — Build a DAG

```{r build-dag}
dag <- emptyOmopDag(
  name        = "NSAID GI Bleed Study",
  version     = "0.1.0",
  description = "Effect of NSAIDs on upper GI bleed risk",
  author      = "Smith J"
)

dag <- addNode(dag, "Age",       conceptId = 4216316L, domain = "Observation")
dag <- addNode(dag, "Celecoxib", conceptId = 1118084L, domain = "Drug")
dag <- addNode(dag, "GI_Bleed",  conceptId = 192671L,  domain = "Condition")
dag <- addNode(dag, "Selected",  conceptId = 4188539L, domain = "Observation")

dag <- addCausal(dag, "Age",       c("Celecoxib", "GI_Bleed"))
dag <- addCausal(dag, "Celecoxib", "GI_Bleed")
dag <- addCausal(dag, "Celecoxib", "Selected")
dag <- addCausal(dag, "GI_Bleed",  "Selected")

dag
```

## 2 — Assign causal roles

```{r set-roles}
dag <- setExposure(dag, "Celecoxib")
dag <- setOutcome(dag,  "GI_Bleed")
dag <- setSelected(dag, "Selected")
```

## 3 — Compute and record adjustment sets

```{r adjustment-sets}
sets <- getMinimalAdjustmentSet(dag)
print(sets)

dag <- setAdjustmentSet(dag, "Celecoxib", "GI_Bleed", index = 1L)
```

## 4 — Check selection bias

```{r selection-bias}
sb <- checkSelectionBias(dag)
print(sb)
```

## 5 — Validate the DAG

```{r validate}
v <- validateDag(dag)
v$overall
lapply(v$checks, function(c) c$message)
```

## 6 — Bind phenotypes

```{r bind-phenotypes}
# Embedded ATLAS JSON
atlas_json <- '{"ConceptSets":[],"PrimaryCriteria":{"CriteriaList":[]}}'
dag <- bindPhenotype(dag, "Celecoxib", type = "atlasJson", definition = atlas_json)
dag <- bindPhenotype(dag, "GI_Bleed",  type = "atlasJson", definition = atlas_json)

# Demographic metadata (no cohort generated)
dag <- bindPhenotype(dag, "Age", type = "demographics", definition = "year_of_birth")

# PhenotypeLibrary reference (resolved before entering a locked environment)
commit <- paste(rep("a", 40), collapse = "")
dag <- bindPhenotype(dag, "Celecoxib",
  type       = "PhenotypeLibrary",
  definition = list(phenotypeId = 1L, commitHash = commit),
  alias      = "pl_ref"
)

# Switch active binding
dag <- setActiveBinding(dag, "Celecoxib", alias = "default")
```

## 7 — Resolve PhenotypeLibrary bindings

```{r resolve, eval = FALSE}
# Run on an internet-connected machine before entering a locked hospital environment
dag <- resolvePhenotypes(dag)
```

## 8 — Visualise

```{r plot-dag}
p <- plotDag(dag)
print(p)
```

## 9 — Execute cohorts

```{r execute, eval = FALSE}
cd <- DatabaseConnector::createConnectionDetails(
  dbms     = "postgresql",
  server   = "my-cdm-server/cdm_v54",
  user     = "analyst",
  password = rstudioapi::askForPassword()
)

result <- executeDagPhenotypes(
  dag,
  connectionDetails = cd,
  cdmSchema         = "cdm_v54",
  resultsSchema     = "results",
  cohortTable       = "cohort"
)
print(result)
```

## 10 — Export and re-import

```{r roundtrip, eval = FALSE}
exportOmopDag(dag, "nsaid_gi_bleed_v1.json")

dag2 <- importOmopDag("nsaid_gi_bleed_v1.json")
dag2
```
